Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-03 foreground_delay_spectrum_modeling.tex
From v0.2: =========
(*) Edited second half of the foreground_power_spectra_modeling.tex file beyond section 6.
(*) Removed unnecessary .eps files from the version and use only those needed to
compile.
(*) Simplified justification of agreement of data with simulations.

We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using a baseline (antenna pair) based ``delay spectrum'' analysis technique. We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies. % Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution. 
We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission. By tying model parameters to those of a subset of first season of observing with the MWA, the ratio of our predictions to observations, subject to expected errors, are consistent with unity at $\gtrsim$~90\% confidence level. A wedge-shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed. Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge-heavy features in the wedge even on large antenna spacings, hitherto not predicted. Features in the inner regions of the wedge are predominantly due to compact foreground objects. This results in a ``pitchfork'' shaped delay spectrum. We also provide a practical diagnostic tool for data analysis that helps in identifying interferometers with high foreground contamination.

@@

In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by an interferometer, as seen from equation~\ref{eqn:delay-transform}. It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes. For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$. For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.

@@

Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east and north (along local meridian) directions at the MWA site. For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively. For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set. Signals from this direction are recorded with a negative delay on interferometers whose antenna spacings are oriented towards the east. Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward oriented antenna spacings.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with MWA data for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon limit and antenna spacing. The features in the wedge appear to be brighter and the spillover contamination appears to be higher in the {\it off--zenith} pointing relative to the {\it zenith} pointing due to the presence of brighter foregrounds such as the Galactic center and the Galactic plane. The logarithmic color scale (shown at the right) is common to both panels.\label{fig:fhd_data}}
\end{figure}

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly. This is predominant at short baselines. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we note that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool. Hence, our discussion and conclusions still hold despite imperfections in deconvolution.

@@

The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is analytically modeled based on a 4$\times$4 phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane. We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns. The models of power pattern used in data analysis take into account effects of mutual coupling of dipoles in the tile besides finite ground plane effects. Our models were found to be in good agreement with those used in the data analysis.

@@

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their similar flux sensitivity and angular resolution, and complimentary survey footprints covering the entire sky. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for discrete sources.

From the SUMSS catalog, we select compact sources whose deconvolved major axes are equal to 0. From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint. Compact sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey. Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$. From this subset, we choose compact objects with $S_{185}\geq 10$~Jy. The selection of such bright compact objects is not affected by minor differences in sensitivity of the two surveys. We verified that our selection criteria ensure a similar areal density of sources in the two surveys.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon limit and antenna spacing. The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. The axes and color scale are identical to those in figure~\ref{fig:fhd_data}. \label{fig:sim_data}}
\end{figure}

In order to make a quantitiative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs. We assume the same holds true for our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}. Besides, delay spectra from simulations and data each have uncertainties arising from thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim 1.4\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise. We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|/|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively. Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.

Figure~\ref{fig:data-sim-ratio-snr} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}. The probability densities and cumulative distributions are shown in black and gray respectively. Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$. The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

These uncertainties are presented only to confirm the qualitative agreement already seen between figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved. However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative. Accounting for such uncertainties will only improve our confidence level in this comparison. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground window}.

@@

\item {\it Sky Model}: Our model of the sky consists of bright compact sources and diffuse emission. Both of these kinds of emission are anisotropic and so is the resulting composite model. In fact, the pointing for MWA observations is chosen from regions of low foreground emission. In our study, we have divided the transiting sky into four different ``bow-tie'' shaped regions with equal areas and zenith as the origin. This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer baselines by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East. We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$. The bin centers are oriented towards South-East, East, North-East, and North respectively.

\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are phased electronically, the power pattern of the tile changes in any observing mode that tracks the source. Our observations consist of a combination of allowing the sky to drift and tracking the sky. In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum. The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours (see figure~\ref{fig:pointings}). For comparisons, we choose two snapshot observations at LST of 22.08~hours (off--zenith pointing) and 0.09~hours (zenith pointing).

\item {\it Instrumental Bandpass}: The instrumental bandpass characteristics have a direct effect on the delay transform owing to a Fourier relationship between the two. Due to excision of noisy edge channels repeated in each coarse channel of our bandpass, we see repeated patterns of the {\it foreground wedge} in the delay spectra. As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects. Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.

@@

The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors. Due to numerous combinations of such factors, we use unique codes for these combinations as described in table~\ref{tab:codes}. The first column refers to the place holder of the specific parameter in the code sequence from left to right. The parameter codes are specified in this order. The second column gives the parameter used in describing features in delay spectrum. The third column denotes the code describing the parameter. Snapshot index uniquely specifies the snapshot and carries a numeric code (1 and 2 for off-zenith and on-zenith snapshots respectively) leading the code sequence. Following this, the type of emission is specified as compact, diffuse, Galactic center, or Galactic plane using codes C, D, GC, or GP respectively. Although in reality, a feature could be due to a combination of emission types, the code we use refers to the predominant type of emission that can be attributed to the feature. The next code specifies the direction of this predominant source of emission. The code values for the direction of emission are the usual abbreviations of the cardinal and ordinal directions besides Z (zenith). Following this is the code for the baseline orientation bin in the half-plane specified as S, SE, E, and NE along directions south, south-east, east, and north-east respectively. Code `A' is used to refer to all baselines. Finally, the parameter describing the region of antenna power pattern coincident with the predominant emission responsible for the feature is specified as P (primary lobe) and S (secondary lobe). The secondary lobe code carries a numeral suffix to further specify uniquely the sidelobe in the direction of emission. For example, a feature code 1-D-W-E-P indicates it is from the first snapshot (1) arising due to diffuse emission (D) westward (W) of zenith observed by predominantly eastward (E) baselines centered on the primary beam (P) of the antenna power pattern. Simiarly, feature code 2-GC-W-NE-S3 implies the feature is seen in the second snapshot (2) due to the Galactic Center (GC) that is westward (W) of zenith by north-eastward (NE) baselines and the region coincides with the third sidelobe (S3) in the direction of emission.

\begin{table*}[htp]
\caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
\centering
\begin{tabular}{ccc}
\tableline\tableline\tableline
Order & Parameter Description & Code \\
\tableline\tableline
1 & Snapshot index & 1, 2, \ldots \\
2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
& & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
& (in the direction of emission) & \\
\tableline
\end{tabular}
\end{table*}

We wish to note that numerous features overlap at varying levels of significance as a result of various combiantions of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.

@@

Figure~\ref{fig:csm} shows the noiseless delay spectra obtained from compact foreground objects. The off-zenith and on-zenith snapshot delay spectra are shown in figures~\ref{fig:csm_delay_spectrum_snap0} and \ref{fig:csm_delay_spectrum_snap1} respectively.

>>>>

We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using an antenna pair based ``delay spectrum'' analysis technique. We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies. % Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution. 
We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse and compact Galactic and extragalactic emission. By tying model parameters to those of a subset of first season of observing with the MWA, our simulations agree with the observed data to within expected uncertainty. A wedge--shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed. Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge--heavy {\it two--pronged fork}--like features even on wide antenna spacings, hitherto not predicted. Features in the inner regions of the wedge are predominantly due to compact foreground objects. This results in a {\it pitchfork}--shaped delay spectrum. The bright Galactic center severly contaminates the {\it EoR window} on certain antenna spacings in certain observations. For such cases, we present an effective practical tool for screening interferometers with high foreground contamination and mitigating their effects on EoR data analysis.

@@

In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by an interferometer, as seen from equation~\ref{eqn:delay-transform}. It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales. For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$. For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.

@@

Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the MWA site. For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively. For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set. Signals from this direction are recorded with a negative delay on interferometers whose antenna spacings are oriented towards the east. Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward oriented antenna spacings.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with MWA data for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The features in the wedge appear to be brighter and the spillover contamination appears to be higher in the {\it off--zenith} pointing relative to the {\it zenith} pointing due to the presence of brighter foregrounds such as the Galactic center and the Galactic plane. The logarithmic color scale (shown at the right) is common to both panels.\label{fig:fhd_data}}
\end{figure}

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly. This is predominant at short antenna spacings. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool. Hence, our discussion and conclusions are valid despite imperfections in deconvolution.

@@

The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is analytically modeled based on a 4$\times$4 phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane. We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns. The models of power pattern used in data analysis take into account effects of mutual coupling of dipoles in the tile besides finite ground plane effects. Our models were found to be in reasonable agreement with those used in the data analysis.

@@

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their matched flux sensitivity and angular resolution, and complimentary survey footprints covering the entire sky. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for discrete sources.

From the SUMSS catalog, we select compact sources whose deconvolved major axes are equal to 0\arcsec. From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint. Compact sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey. Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$. From this subset, we choose compact objects with $S_{185}\geq 10$~Jy. The selection of such bright compact objects is not affected by minor differences in sensitivity of the two surveys. We verified that our selection criteria ensure a similar areal density of sources in the two surveys.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. The axes and color scale are identical to those in figure~\ref{fig:fhd_data}. \label{fig:sim_data}}
\end{figure}

In order to make a quantitiative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$0.7 and $\sim$0.3 respectively. In addition, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise. We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively. % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.

% Figure~\ref{fig:data-sim-ratio} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}. The probability densities and cumulative distributions are shown in black and gray respectively. Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$. The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.

Figure~\ref{fig:data-sim-ratio} shows histograms of distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with aforementioned uncertainties in foreground models and thermal noise in measurements. Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz. Availability of an accurate foreground model will improve the agreement of simulations with observed data.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

These uncertainties are presented only to confirm the qualitative agreement already seen between figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved. However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative. Accounting for such uncertainties will only improve our confidence level in this comparison. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

\item {\it Sky Model}: Our model of the sky consists of bright compact sources and diffuse emission on diverse spatial scales as seen from figures~\ref{fig:DSM} and \ref{fig:CSM}. It also consists of localized regions of strong emission such as the Galactic plane and Galactic center. The resulting sky emission is anisotropic. In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission. In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions with equal areas around zenith as the origin. This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.

\item {\it Interferometer Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer antenna spacings by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East. We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$. The bin centers are oriented towards South-East, East, North-East, and North respectively. Since the delays registered at an interferometer depend on the orientation of its antenna spacing vector, binning by interferometer orientation allows us to attribute the delay spectrum to different sky directions.

\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the power pattern of the tile changes in any observing mode that tracks the source. Our observations consist of a combination of allowing the sky to drift and tracking the sky. In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum (figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in figure~\ref{fig:pointings}.

\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to a Fourier relationship between the two. Due to excision of noisy and aliased edge channels repeated in each coarse channel of our passband, we saw repeated patterns of the {\it foreground wedge} in the delay spectra. As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects. Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.

@@

The foreground signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors. Due to numerous combinations of such factors, we use unique annotations to characterize these signatures.

% The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors. Due to numerous combinations of such factors, we use unique codes for these combinations as described in table~\ref{tab:codes}. The first column refers to the place holder of the specific parameter in the code sequence from left to right. The parameter codes are specified in this order. The second column gives the parameter used in describing features in delay spectrum. The third column denotes the code describing the parameter. Snapshot index uniquely specifies the snapshot and carries a numeric code (1 and 2 for off-zenith and on-zenith snapshots respectively) leading the code sequence. Following this, the type of emission is specified as compact, diffuse, Galactic center, or Galactic plane using codes C, D, GC, or GP respectively. Although in reality, a feature could be due to a combination of emission types, the code we use refers to the predominant type of emission that can be attributed to the feature. The next code specifies the direction of this predominant source of emission. The code values for the direction of emission are the usual abbreviations of the cardinal and ordinal directions besides Z (zenith). Following this is the code for the baseline orientation bin in the half-plane specified as S, SE, E, and NE along directions south, south-east, east, and north-east respectively. Code `A' is used to refer to all baselines. Finally, the parameter describing the region of antenna power pattern coincident with the predominant emission responsible for the feature is specified as P (primary lobe) and S (secondary lobe). The secondary lobe code carries a numeral suffix to further specify uniquely the sidelobe in the direction of emission. For example, a feature code 1-D-W-E-P indicates it is from the first snapshot (1) arising due to diffuse emission (D) westward (W) of zenith observed by predominantly eastward (E) baselines centered on the primary beam (P) of the antenna power pattern. Simiarly, feature code 2-GC-W-NE-S3 implies the feature is seen in the second snapshot (2) due to the Galactic Center (GC) that is westward (W) of zenith by north-eastward (NE) baselines and the region coincides with the third sidelobe (S3) in the direction of emission.

% \begin{table*}[htp]
% \caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
% \centering
% \begin{tabular}{ccc}
% \tableline\tableline\tableline
% Order & Parameter Description & Code \\
% \tableline\tableline
% 1 & Snapshot index & 1, 2, \ldots \\
% 2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
% 3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
% & & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
% 4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
% 5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
% & (in the direction of emission) & \\
% \tableline
% \end{tabular}
% \end{table*}

It may be noted that numerous features overlap at varying levels of significance as a result of various combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as interferometer orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.

@@

Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum arising out of our compact objects sky model for {\it off--zenith} and {\it zenith} pointings (shown in figure~\ref{fig:CSM}) in the top and bottom panels respectively.
